PREVIC: An adaptive parent report measure of expressive vocabulary in children between 3 and 8 years of age

### 2025-02-20 paper/manuscript.tex
update paper link

\linenumbers
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={PREVIC: An adaptive parent report measure of expressive vocabulary in children between 3 and 8 years of age},
  pdfauthor={Manuel Bohn1,2, Julia Prein2, Jonas Engicht3, Daniel Haun2, Natalia Gagarina4, \& Tobias Koch3},
  pdflang={en-EN},
  pdfkeywords={language development, vocabulary, individual differences, Item Response Models},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{PREVIC: An adaptive parent report measure of expressive vocabulary in children between 3 and 8 years of age}
\author{Manuel Bohn\textsuperscript{1,2}, Julia Prein\textsuperscript{2}, Jonas Engicht\textsuperscript{3}, Daniel Haun\textsuperscript{2}, Natalia Gagarina\textsuperscript{4}, \& Tobias Koch\textsuperscript{3}}
\date{}

@@

Learning language is one of the key developmental objectives for children. This learning process is highly variable and leads to persistent individual differences which are related to a wide range of outcome measures later in life (Bleses, Makransky, Dale, Højen, \& Ari, 2016; Bornstein, Hahn, Putnick, \& Pearson, 2018; Golinkoff, Hoff, Rowe, Tamis-LeMonda, \& Hirsh-Pasek, 2019; Marchman \& Fernald, 2008; Morgan, Farkas, Hillemeier, Hammer, \& Maczuga, 2015; Pace, Alper, Burchinal, Golinkoff, \& Hirsh-Pasek, 2019; Pace, Luo, Hirsh-Pasek, \& Golinkoff, 2017; Schoon, Parsons, Rush, \& Law, 2010; Walker, Greenwood, Hart, \& Carta, 1994). For example, in a longitudinal study spanning 29 years, Schoon et al. (2010) found that relatively poorer language skills at age five were associated with lower levels of mental health at age 34. Given the high predictive validity of early language abilities, researchers and practitioners alike need high-quality, easy access measures to assess individual differences. However, such measures are rare and those that exist often come with substantial licensing costs. In this paper, we describe the development of an open, efficient and valid measure of individual differences in expressive vocabulary.

@@

However, the use of the CDI -- in typically developing children -- is limited to 37 months of age. Beyond this point, most children are reported to say all the words on the list. Consequently, there is a need for a comparable measure that can be applied to older children. Existing instruments focusing on general cognitive development often include language scales; however, these scales lack detailed information and fail to capture individual differences effectively (Ireton \& Glascoe, 1995). For example, the Ages and Stages Questionnaire (ASQ) at 36 months comprises only six items that encompass general communicative behavior, such as whether the child can say their full name when prompted (Squires, Bricker, Twombly, et al., 2009). One notable example of a dedicated language measure for older children is the Developmental Vocabulary Assessment for Parents (DVAP, Libertus, Odic, Feigenson, \& Halberda, 2015). The DVAP is derived from the words used in the Peabody Picture Vocabulary Test (PPVT, Dunn \& Dunn, 1965), a widely used direct measure of receptive vocabulary. As perhaps expected, the DVAP demonstrates high convergent validity, as evidenced by its strong correlation with the PPVT. However, the proprietary nature of the PPVT limits the utility of the DVAP for researchers.\footnote{When the first author approached the license holder of the PPVT in Germany to ask if we could use the German version of the PPVT to build a parental report measure, we were told that we would have to pay for every administration of the new measure and we would not be allowed to openly share the materials.} As a consequence, it is unlikely that a comparable ``success story'' -- as observed with the CDI -- will emerge where researchers have adapted the original English form to different languages and more efficient forms.

A more general issue with existing language measures -- including PPVT and DVAP -- is a lack of psychometric grounding. Items that make up the scale are selected based on researchers intuitions and there is no clear measurement model that explicates how the different items and test scores are linked to the construct in question (Borsboom, 2006). Item response theory (IRT) offers a theoretical framework to fill this gap and provides a toolkit to develop tasks with a solid psychometric foundation (Kubinger, 2006; Lord, 2012). Within IRT, all items are assumed to measure the same latent construct. Each item is linked to the construct by a probability function (e.g.~a logistic curve) which determines how likely a particular response is for individuals with different values on the latent construct. The shape of this curve is defined by the difficulty of an item (i.e.~the value of the latent construct when the probability to solve the item is 50\%) and its discrimination (i.e.~the slope of the curve showing how the probability to solve the item changes with increasing levels on the latent construct). In a \emph{Rasch} model (or one-parameter IRT model), all items are assumed to have equal item discriminations, resulting in parallel item characteristic curves. The great benefit of IRT is that models are testable in that we can quantify the fit of each model and compare competing models. For each item, we can compute fit statistics that indicate how well the model captures the response pattern to the item. Test construction is straightforward in this framework; items are selected that improve the fit to the model.

Besides many other advantages (e.g., objective specificity, sum scores can be used as sufficient statistics), the Rasch model also allows for adaptive testing. All items in a test that conforms to the Rasch model measure the same latent ability and all that differs is their difficulty. As a consequence, the ability of a person can be estimated independent of the particular items that have been completed. This characteristic is leveraged during adaptive testing: individuals do not simply respond to all items in the task but only to the items that are optimally informative given their -- constantly updated -- individual value on the latent construct. Because all items measure the same latent ability, the resulting scores are nevertheless comparable.

The downside of IRT-based test construction -- and probably the reason it is not used more often -- is that it requires a larger initial investment. To be able to remove items with a poor fit during the selection process requires an initial item pool that is substantially larger than the desired size of the final task. Adaptive testing also needs a large item pool so that there is a sufficient number of items that are optimally informative for different regions of the latent construct. Furthermore, to obtain solid estimates for the item parameters it takes large sample sizes. Yet, we believe that these initial costs are clearly outweighed by the benefits that come with IRT-based test construction in the long run.

@@

Our goal was to develop a high-quality and easy-access vocabulary checklist beyond the CDI for children between three and eight years of age. To ensure the psychometric quality of the task and to allow for adaptive testing, we used IRT to guide item selection and the construction of the item pool. In particular, we constructed the item pool so that all items fit the simplest version of an IRT model, the one-parameter logistic (1PL) or Rasch model (Rasch, 1980). The main reason behind this fairly restrictive approach was that only when the Rasch model holds is the number of solved items (sum score) a sufficient statistic and can be used to represent an individual's value on the latent construct (Birnbaum, 1986). To ensure easy-access, we implemented the checklist as an interactive web-app. Furthermore, the task, the item pool and all associated materials are openly available for other researchers to use. In what follows, we describe the implementation of the task, the construction of the item pool and the item selection procedure. Finally, we report the results on a study assessing the convergent validity of a task that comprises all items in the pool.

@@

\hypertarget{task-design-an-implementation}{%
\subsection{Task design an implementation}\label{task-design-an-implementation}}

We decided to use an interactive format instead of presenting parents with a long list of words in order to increase number of items while keeping the task engaging. The task was implemented as a web-app using \texttt{html} and \texttt{JavaScript} and ran in every modern web-browser on computers, tablets and smartphones. Words were presented one-by-one and caregivers could indicate whether or not their child says a word either by using the familiar swipe-left/swipe-right functionality, by clicking symbols or using arrow-keys on a keyboard (see Figure \ref{fig:fig1}A). For example, caregivers saw the word ``Jacke'' (en: jacket) on a card (color-coded by part of speech: noun = blue, adjective = orange, verb = green) in the center of the screen; to report that their child says the word, they would swipe the card to the right side of the screen which would make the card go away and the next in the deck appear. We included a lightweight back-end that registered the last completed trial so that caregivers could take breaks and even switch devices during the task.

@@

Our goal was to create an item pool with items of different word classes and varying semantical difficulty. We used age-of-acquisition (AoA) ratings as a rough indicator of anticipated item difficulty. Previous work has shown strong associations between AoA ratings and how likely children are to know a word (Bohn et al., 2023; Bohn, Tessler, Merrick, \& Frank, 2021). We started the process by compiling a list with AoA ratings for 3,921 German words from various sources (Birchenough, Davies, \& Connelly, 2017; Łuniewska et al., 2019; Schröder, Gemballa, Ruppin, \& Wartenburger, 2012). We excluded words with AoA ratings above ten. The remaining words were ordered by rated AoA and then split into ten lists with 344 words each. A research assistant with a background in linguistics went through the lists and selected words that a) were indicative of language abilities more broadly (avoiding very specialized terms) and b) that were different from one another in that they were semantically unrelated (to avoid words that are learned in the same context). For each list, we aimed for roughly 35 words from the three word classes; 17 nouns, nine verbs and nine adjectives. The so-generated item pool had 379 words, of which 197 were nouns, 92 were verbs and 90 were adjectives. Figure \ref{fig:fig2}A shows how the items were distributed across AoA ratings and word types.

@@

Figure \ref{fig:fig2} visualizes the results. On an item level, we saw strong negative correlations between caregiver's responses and rated ages-of-acquisition. The less likely caregiver's were to say their child says a word, the higher was the rated AoA. This relation was the same for nouns, verbs and adjectives. On a child-level, we saw that the older children were, the more words they used according to their caregiver's. These results reflect highly expected patterns and served as a sanity check for the design and implementation of the task.

@@

\caption{Initial item pool. (A) Distribution of items across word types and rated age-of-acquisition. (B) Item-based correlation between rated age-of-acquisition and caregiver responses by word type averaged across participants. (C) Child-based correlation between age and caregiver responses by word type averaged across items.}\label{fig:fig2}
\end{figure}

@@

IRT-models were implemented in a Bayesian framework in \texttt{R} using the \texttt{brms} package (Bürkner, 2017, 2019) unless otherwise stated. We predicted the probability of a correct answer based on a participant's latent ability and item characteristics. We fit two classes of models: Rasch (1PL) and Birnbaum (2PL) models. The main difference between these two models lies in their assumption about how the probability of solving an item changes with ability levels (item discrimination). Here, the Rasch model assumes that the rate of change (i.e.~the slope of the logistic curve) is the same for all items while the 2PL model allows item discrimination parameters to vary between items.

@@

We fit a Rasch Model to the data and computed In- and Outfit values based on draws from the model's expected posterior predictive distribution (using the function \texttt{add\_epred\_draws}) for each item and person combination. In- and Outfit quantify the deviation of a person's response to an item from what the model predicts based on the difficulty of an item and the person's ability parameter. The result was a distribution of values for each item and index. For each item, we then computed the mode for each index. The closer to 1 the index is, the better the fit. Figure \ref{fig:fig3} visualizes the results. We used the cut-off values suggested in the literature (Bond \& Fox, 2013; Debelak, Strobl, \& Zeigenfuse, 2022) and excluded items with In- or Outfit values below 0.7 and above 1.3. Like all heuristics, these cut-offs are to some extent arbitrary. Yet, in the present context they served the purpose of removing a large number of potentially unsuitable items. This procedure led us to exclude 167 of the 379 items in the pool, leaving 212 for the automated item selection.

@@

First we defined an objective function that reflected the selection criteria which would later be used in the automated selection process. Items should vary in their difficulty; we quantified this requirement as the standard deviation of the distance (in difficulty estimates) between adjacent items. Lower values indicate smaller distances and thus an overall more equal spacing. Items should also fit the Rasch model; we quantified this requirement in three ways. First and second, we used the In- and Outfit values for each item computed in the previous step. Third, we computed modification indices for each item. For this, we re-fitted the Rasch model using the package \texttt{lavaan} (Rosseel, 2012) and used the function \texttt{modindices} to obtain modification indices. Broadly speaking, modification indices quantify the improvement in model fit (in terms of the chi-square test statistic) when an item would be dropped.

@@

\hypertarget{psychometric-properties-of-the-final-item-pool}{%
\subsection{Psychometric properties of the final item pool}\label{psychometric-properties-of-the-final-item-pool}}

The final item pool consisted of 89 items of varying difficulty that fit the Rasch model (see Figure \ref{fig:fig4}A). Next, we investigated the reliability and convergent validity of a task including the full item pool (PREVIC).

@@

We assessed convergent validity by comparing PREVIC scores to a direct assessment of children's receptive vocabulary using the oREV (Bohn et al., 2023). The oREV asks children to select a picture (out of four) upon hearing a word. It has 22 items which fit the Rasch model. Because the oREV is also available as a web application, we sent out emails to all caregivers who provided complete data in the data collection that led to the construction of the PREVIC (N = 1190) and asked them to have their child complete the oREV. We obtained oREV data from 692 children (337 female, \(m_{age}\) = 5.78, range = 3.02 - 8.00) which corresponds to a response rate of \textasciitilde{} 58\%.

We found a substantial correlation between caregiver's answers to questions about their children's expressive vocabulary in the PREVIC and a direct assessment of children's receptive vocabulary in the oREV (\emph{r} = 0.54; 95\% CI = 0.48 -- 0.59; Figure \ref{fig:fig4}B). Furthermore, when predicting PREVIC scores by oREV scores and age in a binomial model, we found that oREV scores had a large and positive effect (Figure \ref{fig:fig4}C). These results speak to the convergent validity of the PREVIC.

@@

In order to determine the most informative items, the ability of the child has to be estimated during the test. To achieve this, we implemented a maximum likelihood estimator in \texttt{html} and \texttt{TypeScript} (which is compiled to native \texttt{JavaScript}). As a consequence, the adaptive version is still fully portable and can be run in any modern web-browser. The estimated ability ``is the ability value that maximizes the likelihood function \(L(\theta)\)'' (Magis \& Raîche, 2012), given the item response \(y_i\) (either 0 or 1) and the item difficulty \(\alpha_i\) (Eid, 2011; Eid \& Schmidt, 2014):

\[
L(\theta) = \prod_{i = 1}^{p} \frac{\exp^{y_i * (\theta - \alpha_i)}}{1 + \exp^{\theta - \alpha_i}}
\] The maximum likelihood estimation is implemented using a line search algorithm that converges when the maximum of the likelihood distribution has been reached. Based on the estimated ability, the program will then select the next item from the pool so that the difficulty is nearest to the current ability level (Urry, 1970). This procedure is equivalent to selecting items with the maximum information criterion when using a Rasch model (Magis \& Raîche, 2012).

At the beginning of the test, the ability level is set to 0. A person-specific ability estimation is not yet possible using Maximum Likelihood Estimation after the first item and we therefore followed the convention to set the ability level to -10 if the answer was ``no'' and 10 if the answer was ``yes'' (e.g.~implemented in the \texttt{R} package \texttt{catR}, Magis \& Barrada, 2017). The test then continues until a pre-specified level of measurement precision (standard error of the ability estimate) is reached or until all items have been used. Users can set the desired level of measurement precision at the beginning of the test (e.g.~SE of 0.3, 0.4 or 0.5), which again influences its length (larger SE means shorter test). In the end, the user downloads a file containing the final ability level, the SE at which the test stopped, the answered items (including the word itself and its difficulty) and the participant's response pattern.

@@

The PREVIC fills an important gap in the tool kit of researchers studying language development beyond infancy and particular during the preschool years. It complements direct assessments of children's vocabulary by providing an additional perspective on children's vocabulary skills. Parents observe children for extended periods of time and their assessment therefore provides a more aggregated measure. Parental reports are also immune to momentary fluctuations in children's motivation and attention that might influence the results of direct assessments. Nevertheless, parental reports remain indirect measures and are ideally combined with direct assessments whenever feasible. Given that the PREVIC is short -- in particular the adaptive version rarely takes more than five minutes to complete -- it can easily be filled out by parents during a lab visit. Its implementation as a web-app even allows for sending it to families before or after a visit.

At present, the PREVIC is available only in German. However, with inclusivity and broad applicability in mind, we have made the entire source code available. This not only facilitates its adaptation to other languages and allows researchers to use the same user interface. Encouragingly, preliminary feedback indicates that parents find the interface intuitive and user-friendly. The CDI has seen expansive adaptation across various languages (see Frank et al., 2021 for a summary). Such adaptations are usually not complete translations in that some words are removed and others added to capture the linguistic nuances and specificities of each language. However, Łuniewska et al. (2019) found that the order of acquisition (and thus presumably the difficulty) was similar for many words across seven languages. Hence, most items in the current pool could be translated and re-used if the PREVIC were to adapted to different languages. Nevertheless, a comprehensive reassessment of item properties would be highly desirable. For adaptive testing, it would even be mandatory. Taken together, we hope that our commitment to openness will put the PREVIC on a similar trajectory as the CDI.

@@

\leavevmode\vadjust pre{\hypertarget{ref-eid2011}{}}%
Eid, M. (2011). \emph{Testtheorie} (Erscheint: ET unbestimmt). Göttingen: Hogrefe.

@@

>>>>

\linenumbers
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={PREVIC: An adaptive parent report measure of expressive vocabulary in children between 3 and 8 years of age},
  pdfauthor={Manuel Bohn1,2, Julia Prein1,2, Jonas Engicht3, Daniel Haun2, Natalia Gagarina4, \& Tobias Koch3},
  pdflang={en-EN},
  pdfkeywords={language development, vocabulary, individual differences, Item Response Models},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{PREVIC: An adaptive parent report measure of expressive vocabulary in children between 3 and 8 years of age}
\author{Manuel Bohn\textsuperscript{1,2}, Julia Prein\textsuperscript{1,2}, Jonas Engicht\textsuperscript{3}, Daniel Haun\textsuperscript{2}, Natalia Gagarina\textsuperscript{4}, \& Tobias Koch\textsuperscript{3}}
\date{}

@@

Learning a language is one of the key developmental objectives for children. This learning process is highly variable and leads to persistent individual differences which are related to a wide range of outcome measures later in life (Bleses, Makransky, Dale, Højen, \& Ari, 2016; Bornstein, Hahn, Putnick, \& Pearson, 2018; Golinkoff, Hoff, Rowe, Tamis-LeMonda, \& Hirsh-Pasek, 2019; Marchman \& Fernald, 2008; Morgan, Farkas, Hillemeier, Hammer, \& Maczuga, 2015; Pace, Alper, Burchinal, Golinkoff, \& Hirsh-Pasek, 2019; Pace, Luo, Hirsh-Pasek, \& Golinkoff, 2017; Schoon, Parsons, Rush, \& Law, 2010; Walker, Greenwood, Hart, \& Carta, 1994). For example, in a longitudinal study spanning 29 years, Schoon et al. (2010) found that relatively poorer language skills at age five were associated with lower levels of mental health at age 34. Given the high predictive validity of early language abilities, researchers and practitioners alike need high-quality, easy access measures to assess individual differences. However, such measures are rare and those that exist often come with substantial licensing costs. In this paper, we describe the development of an open, efficient and valid measure of individual differences in expressive vocabulary.

@@

However, the use of the CDI -- in typically developing children -- is limited to 37 months of age. Beyond this point, most children are reported to say all the words on the list. Consequently, there is a need for a comparable measure that can be applied to older children. Even though a wide range of direct language measures exist for preschool and school-aged children, parent report measures can be useful. First, they offer a complementary and perhaps more holistic perspective on children's language abilities because parents rely on their extensive experience with their children when filling out. Second, they are less dependent on situational factors like children's fatigue or shyness compared to direct assessments. Finally, they are easier and more economical to apply because they need less time and do not require trained experimenters. This makes them very valuable research and -- if normed -- screening tools, in particular when dealing with large sample sizes. Existing parent report measures focusing on general cognitive development often include language scales; however, these scales lack detailed information and fail to capture individual differences effectively (Ireton \& Glascoe, 1995). For example, the Ages and Stages Questionnaire at 36 months comprises only six items that encompass general communicative behavior, such as whether the child can say their full name when prompted (Squires, Bricker, Twombly, et al., 2009). One notable example of a dedicated language measure for older children is the Developmental Vocabulary Assessment for Parents (DVAP, Libertus, Odic, Feigenson, \& Halberda, 2015). The DVAP is derived from the words used in the Peabody Picture Vocabulary Test (PPVT, Dunn \& Dunn, 1965), a widely used direct measure of receptive vocabulary. As perhaps expected, the DVAP demonstrates high convergent validity, as evidenced by its strong correlation with the PPVT. However, the proprietary nature of the PPVT limits the utility of the DVAP for researchers.\footnote{When the first author approached the license holder of the PPVT in Germany to ask if we could use the German version of the PPVT to build a parental report measure, we were told that we would have to pay for every administration of the new measure and we would not be allowed to openly share the materials.} As a consequence, it is unlikely that a comparable ``success story'' -- as observed with the CDI -- will emerge where researchers have adapted the original English form to different languages and more efficient forms.

A more general issue with existing language measures -- including PPVT and DVAP -- is a lack of psychometric grounding. Items that make up the scale are selected based on researchers' intuitions and there is no clear measurement model that explicates how the different items and test scores are linked to the construct in question (Borsboom, 2006). Item response theory (IRT) offers a theoretical framework to fill this gap and provides a toolkit to develop tasks with a solid psychometric foundation (Kubinger, 2006; Lord, 2012). In unidimensional IRT models, it is assumed that all items measure the same latent construct. Each item is linked to the construct by a probability function (e.g.~a logistic curve) which determines how likely a particular response is for individuals with different values on the latent ability (see Figure \ref{fig:figr}). The location and shape of this curve is defined by the difficulty of an item (i.e.~the value of the latent construct when the probability to solve the item is 50\%) and its discrimination (i.e.~the slope of the curve showing how the probability to solve the item changes with increasing levels on the latent construct). In a \emph{Rasch} model, all items are assumed to have equal item discriminations, resulting in parallel item characteristic curves (Rasch, 1980). The great benefit of IRT is that models are testable in that we can quantify the fit of each model and compare competing models. For each item, we can compute fit statistics that indicate how well the model captures the response pattern to the item. Test construction is straightforward in this framework; items are selected that improve the fit to the model.

{\centering \includegraphics[width=0.5\linewidth]{../graphs/figr}

\caption{Task implementation. (A) Instructions provided to parents demonstrating the functionality of the task. The word was presented on a card in the middle of the screen. Parents could indicate whether or not their child says a word by swiping the card left (no) or right (yes), touching or clicking the ``yes'' or ``no'' symbol or pressing the left or right arrow key on the keyboard. (B) Screenshot from the task presenting the German word Jacke (en: jacket).}\label{fig:figr}
\end{figure}

Besides many other advantages (e.g., objective specificity, sum scores can be used as sufficient statistics), the Rasch model also allows for adaptive testing. All items in a test that conforms to the Rasch model measure the same latent ability and all that differs is their difficulty. As a consequence, the ability of a person can be estimated independently of the particular items that have been completed. This characteristic is leveraged during adaptive testing: individuals do not simply respond to all items in the task but only to the items that are optimally informative given their -- constantly updated -- individual value on the latent construct. Because all items measure the same latent ability, the resulting scores are nevertheless comparable.

The downside of IRT-based test construction -- and probably the reason it is not used more often -- is that it requires a larger initial investment (Frey, 2020). To be able to remove items with a poor fit during the selection process requires an initial item pool that is substantially larger than the desired size of the final task. Adaptive testing also needs a large item pool so that there is a sufficient number of items that are optimally informative for different regions of the latent construct. Furthermore, to obtain solid estimates for the item parameters it takes large sample sizes. Yet, we believe that these initial costs are clearly outweighed by the benefits that come with IRT-based test construction in the long run.

@@

Our goal was to develop a high-quality and easy-access vocabulary checklist beyond the CDI for children between three and eight years of age. To ensure the psychometric quality of the task and to allow for adaptive testing, we used IRT to guide item selection and the construction of the item pool. We compiled a large initial pool of candidate items. Next, we collected a large data set and analyzed it using the simplest version of an IRT model, the Rasch model (Rasch, 1980). The main reason behind this fairly restrictive approach was that only when the Rasch model holds is the number of solved items (sum score) a sufficient statistic and can be used to represent an individual's value on the latent construct (Birnbaum, 1986). Based on the first analysis, we computed a range of item-level indices that captured how difficult the item was and how well it fit the Rasch model. We then used an automated procedure to construct a smaller pool of items with varying difficulties that all fit the Rasch model. Finally, we report the results of two studies assessing the convergent validity of the task. To ensure easy-access, we implemented the checklist as an interactive web-app. Furthermore, the task, the item pool and all associated materials are openly available for other researchers to use.

@@

\hypertarget{task-design-and-implementation}{%
\subsection{Task design and implementation}\label{task-design-and-implementation}}

We decided to use an interactive format instead of presenting parents with a long list of words in order to increase the number of items while keeping the task engaging. The task was implemented as a web-app using \texttt{html} and \texttt{JavaScript} and ran in every modern web-browser on computers, tablets and smartphones. Words were presented one-by-one and caregivers could indicate whether or not their child says a word either by using the familiar swipe-left/swipe-right functionality, by clicking symbols or using arrow-keys on a keyboard (see Figure \ref{fig:fig1}A). For example, caregivers saw the word ``Jacke'' (en: jacket) on a card (color-coded by part of speech: noun = blue, adjective = orange, verb = green) in the center of the screen; to report that their child says the word, they would swipe the card to the right side of the screen which would make the card go away and the next in the deck appear. We included a lightweight back-end that registered the last completed trial so that caregivers could take breaks and even switch devices during the task. There was no time limit for the completion of the task.

@@

Our goal was to create an item pool with items of different word classes and varying semantical difficulty. We used Age-of-Acquisition (AoA) ratings as a rough indicator of anticipated item difficulty. Previous work has shown strong associations between AoA ratings and how likely children are to know a word (Bohn et al., 2023; Bohn, Tessler, Merrick, \& Frank, 2021). We started the process by compiling a list with AoA ratings for 3,921 German words from various sources (Birchenough, Davies, \& Connelly, 2017; Łuniewska et al., 2019; Schröder, Gemballa, Ruppin, \& Wartenburger, 2012). We excluded words with AoA ratings above ten. The remaining words were ordered by rated AoA and then split into ten lists with 344 words each. A research assistant with a background in linguistics went through the lists and selected words that a) were indicative of language abilities more broadly (avoiding very specialized terms) and b) that were different from one another in that they were semantically unrelated (to avoid words that are learned in the same context). For each list, we aimed for roughly 35 words from the three word classes; 17 nouns, nine verbs and nine adjectives. The so-generated item pool had 379 words, of which 197 were nouns, 92 were verbs and 90 were adjectives. Figure \ref{fig:fig2}A shows how the items were distributed across AoA ratings and word types.

@@

Figure \ref{fig:fig2} visualizes the results. On an item level, we saw strong negative correlations between caregiver's responses and rated ages-of-acquisition. The less likely caregiver's were to say their child says a word, the higher was the rated AoA. This relation was the same for nouns, verbs and adjectives. On a child-level, we saw that the older children were, the more words they used according to their caregivers. These results reflect highly expected patterns and served as a sanity check for the design and implementation of the task.

@@

\caption{Initial item pool. (A) Distribution of items across word types and rated age-of-acquisition. (B) Item-based association between rated age-of-acquisition and caregiver responses by word type averaged across participants. (C) Child-based association between age and caregiver responses by word type averaged across items.}\label{fig:fig2}
\end{figure}

@@

IRT-models were implemented in a Bayesian framework in \texttt{R} using the \texttt{brms} package (Bürkner, 2017, 2019) unless otherwise stated. We predicted the probability of a correct answer based on a participant's latent ability and item characteristics. We fit two classes of models: Rasch and Birnbaum (2PL) models. The main difference between these two models lies in their assumption about how the probability of solving an item changes with ability levels (item discrimination). Here, the Rasch model assumes that the rate of change (i.e.~the slope of the logistic curve) is the same for all items while the 2PL model allows item discrimination parameters to vary between items.

@@

In- and Outfit quantify the deviation of a person's response to an item from what the model predicts based on the difficulty of an item and the person's ability parameter. That is, both indices are a direct measure of how well the model was able to predict the responses to an item. We fit a Rasch Model to the data and computed In- and Outfit values based on draws from the model's expected posterior predictive distribution (using the function \texttt{add\_epred\_draws}) for each item and person combination. For each draw, we computed the residual between predicted and observed responses. The mean of the squared residuals is the Outfit; to obtain the Infit, the mean squared residuals are weighted by item information. The result was a distribution of values for each item and index. For each item, we then computed the mode for each index. The closer to 1 the index is, the better the fit. Figure \ref{fig:fig3} visualizes the results. We used the cut-off values suggested in the literature (Bond \& Fox, 2013; Debelak, Strobl, \& Zeigenfuse, 2022) and excluded items with In- or Outfit values below 0.7 and above 1.3. Like all heuristics, these cut-offs are to some extent arbitrary. Yet, in the present context they served the purpose of removing a large number of potentially unsuitable items. This procedure led us to exclude 167 of the 379 items in the pool, leaving 212 for the automated item selection.

@@

First we defined an objective function that reflected the selection criteria which would later be used in the automated selection process. Items should vary in their difficulty but still cover all sections of the latent ability; we quantified this requirement as the standard deviation of the distance (in difficulty estimates) between adjacent items. The distance between adjacent items was computed by first sorting all items in the subset by difficulty and then subtracting the difficulty of adjacent items. Lower values indicate smaller distances and thus an overall more equal spacing. Items should also fit the Rasch model; we quantified this requirement in three ways. First and second, we used the In- and Outfit values for each item computed in the previous step. Third, we computed modification indices for each item. For this, we re-fitted the Rasch model using the package \texttt{lavaan} (Rosseel, 2012) and used the function \texttt{modindices} to obtain modification indices. Broadly speaking, modification indices quantify the improvement in model fit (in terms of the chi-square test statistic) when an item would be dropped (Rosseel, 2012).

@@

In order to determine the most informative items, the ability of the child has to be estimated during the test. To achieve this, we implemented a maximum likelihood estimator in \texttt{html} and \texttt{TypeScript} (which is compiled to native \texttt{JavaScript}). As a consequence, the adaptive version is still fully portable and can be run in any modern web-browser. The estimated ability ``is the ability value that maximizes the likelihood function \(L(\theta)\)'' (Magis \& Raîche, 2012), given the item response \(y_i\) (either 0 or 1) and the item difficulty \(\alpha_i\) (Eid \& Schmidt, 2014):

\[
L(\theta) = \prod_{i = 1}^{p} \frac{\exp^{y_i * (\theta - \alpha_i)}}{1 + \exp^{\theta - \alpha_i}}
\] The maximum likelihood estimation is implemented using a line search algorithm that converges when the maximum of the likelihood distribution has been reached. Based on the estimated ability, the task will then select the next item from the pool so that the difficulty is nearest to the current ability level (Urry, 1970). This procedure is equivalent to selecting items with the maximum information criterion when using a Rasch model (Magis \& Raîche, 2012).

At the beginning of the test, the ability level is set to 0. A person-specific ability estimation is not yet possible using Maximum Likelihood Estimation after the first item and we therefore followed the convention to set the ability level to -10 if the answer was ``no'' and 10 if the answer was ``yes'' (e.g.~implemented in the \texttt{R} package \texttt{catR}, Magis \& Barrada, 2017). The test then continues until a pre-specified level of measurement precision (standard error of the ability estimate) is reached or until all items have been used. Users can set the desired level of measurement precision at the beginning of the test (e.g.~SE of 0.3, 0.4 or 0.5), which again influences its length (larger SE means shorter test). In the end, the user downloads a file containing the following information: the estimate of the latent ability of the participant (on the same scale as item difficulties), the SE of the ability estimate (also used to terminate the test), the an the final ability level, the answered items (including the word itself and its difficulty) and the participant's response pattern.

\hypertarget{psychometric-properties}{%
\subsection{Psychometric properties}\label{psychometric-properties}}

The final item pool consisted of 89 items of varying difficulty that fit the Rasch model (see Figure \ref{fig:fig4}A). Next, we investigated the reliability and convergent validity of a task including the full item pool as well as the adaptive version.

@@

We assessed convergent validity in two studies. First, we compared PREVIC scores to a direct assessment of children's receptive vocabulary using the oREV (Bohn et al., 2023). The oREV asks children to select a picture (out of four) upon hearing a word. It has 22 items which fit the Rasch model. Because the oREV is also available as a web application, we sent out emails to all caregivers who provided complete data in the data collection that led to the construction of the PREVIC (N = 1190) and asked them to have their child complete the oREV. We obtained oREV data from 692 children (337 female, \(m_{age}\) = 5.78, range = 3.02 - 8.00) which corresponds to a response rate of \textasciitilde{} 58\%. We found a substantial correlation between caregiver's answers to questions about their children's expressive vocabulary in the PREVIC and a direct assessment of children's receptive vocabulary in the oREV (\emph{r} = 0.54; 95\% CI = 0.48 -- 0.59; Figure \ref{fig:fig4}B).

Second, we directly assessed children's productive vocabulary using the AWST-R (Aktiver Wortschatztest für 3- bis 5-jährige Kinder -- Revision, Kiese-Himmel, 2005). The AWST-R is not available as an online version and we therefore resorted to in person testing. Children were tested in a separate room in a child laboratory by a trained experimenter while parents filled out the adaptive version of the PREVIC on a tablet in the waiting room. We used an SE of 0.4 for the ability estimate as criterion for termination of the adaptive PREVIC. A total of 70 children and their parents participated in the study (38 female, \(m_{age}\) = 5.02, range = 4.12 - 5.94). We found a substantial correlation between PREVIC and AWST-R scores (\emph{r} = 0.36; 95\% CI = 0.13 -- 0.55; Figure \ref{fig:fig4}C). This correlation was lower compared to the oREV even though the AWST-R is -- like the PREVIC -- a measure of expressive vocabulary. We discuss this result in more detail below. Nevertheless, the results of these two studies speak to the convergent validity of the PREVIC

@@

@@

The PREVIC fills an important gap in the tool kit of researchers studying language development beyond infancy and particular during the preschool years. It complements direct assessments of children's vocabulary by providing an additional perspective on children's vocabulary skills. Parents observe children for extended periods of time and their assessment therefore provides a more aggregated measure. Parental reports are also immune to momentary fluctuations in children's motivation and attention that might influence the results of direct assessments. Nevertheless, parental reports remain indirect measures and are ideally combined with direct assessments whenever feasible. Given that the PREVIC is short -- in particular the adaptive version rarely takes more than five minutes to complete -- it can easily be filled out by parents during a lab or any other institutional, e.g.~pediatrician, visit. Its implementation as a web-app even allows for sending it to families before or after a visit.

At present, the PREVIC is available only in German. However, with inclusivity and broad applicability in mind, we have made the entire source code available. This not only facilitates its adaptation to other languages and allows researchers to use the same user interface. Encouragingly, preliminary feedback indicates that parents find the interface intuitive and user-friendly. The CDI has seen expansive adaptation across various languages (see Frank et al., 2021 for a summary). Such adaptations are usually not complete translations in that some words are removed and others added to capture the linguistic nuances and specificities of each language. However, Łuniewska et al. (2019) found that the order of acquisition (and thus presumably the difficulty) was similar for many words across seven languages. Hence, most items in the current pool could be translated and re-used if the PREVIC were to be adapted to different languages. Nevertheless, a comprehensive reassessment of item properties would be highly desirable. For adaptive testing, it would even be mandatory. Taken together, we hope that our commitment to openness will put the PREVIC on a similar trajectory as the CDI.

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-frey2020computerisiertes}{}}%
Frey, A. (2020). Computerisiertes adaptives testen. In H. Moosbrugger \& A. Kelava (Eds.), \emph{Testtheorie und fragebogenkonstruktion} (pp. 501--525). Berlin, Heidelberg: Springer Berlin Heidelberg. \url{https://doi.org/10.1007/978-3-662-61532-4_20}
